---
title: "uc511 - Spatially Balanced Sampling"
author: "Phil Davies"
date: "19/11/2023"

always_allow_html: true

#output: rmarkdown::html_vignette
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    fig_caption: yes

vignette: >
  %\VignetteIndexEntry{uc511 - Spatially Balanced Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(bookdown)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(uc511)
```

# uc511: Spatially balanced sampling.

This vignette is intended to provide details of the functions and methods in the __uc511__ package. The __uc511__ package encapsulates a number of spatially balanced sampling algorithms namely, Balanced acceptance sampling (BAS) (equal, unequal, seed point, panels), Halton frames (for discretizing a continuous resource), HIP (equal probability), Spatial balance measures (Voronoi and Modified Moran's I), Variance estimators (local mean, nearest neighbour and quasi bootstrap), DAS and the Spatial stratification algorithm.

This vignette is divided into the following sections:

## Simple Random Sampling (SRS)

### uc511::SRS()

This function invokes __base::sample()__ to draw a random sample using a user specified random seed. 

The following parameters are supported:

+ __seed__ - The random seed to be used to draw the current sample. The default is 42.
+ __total_rows__ - The total number of rows in our input dataset.
+ __sample_size__ - The number of rows wanted in our random sample.

All parameter values must be numeric and have values greater than zero, __sample_size__ must be less than __total_rows__.

#### uc511::SRS() code example

Function __uc511::SRS()__ returns a list of size __sample_size__.

```{r}
# Create a random sample with a seed of 99.
rand_samp <- uc511::SRS(seed = 99, total_rows = 100, sample_size = 20)
rand_samp
```

### uc511::SRS() from the uc511 Shiny App

The uc511 Shiny App has a menu option (SRS) to invoke Simple Random Sampling.




## Balanced acceptance sampling (BAS)

## Halton Frames (HF)

