---
title: "uc511 - Spatially Balanced Sampling"
author: "Phil Davies"
date: "19/11/2023"

always_allow_html: true

#output: rmarkdown::html_vignette
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    fig_caption: yes

vignette: >
  %\VignetteIndexEntry{uc511 - Spatially Balanced Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(bookdown)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(uc511)
```

# uc511: Spatially balanced sampling.

This vignette is intended to provide details of the functions and methods in the __uc511__ package. The __uc511__ package encapsulates a number of spatially balanced sampling algorithms namely, Balanced acceptance sampling (BAS) (equal, unequal, seed point, panels), Halton frames (for discretizing a continuous resource), HIP (equal probability), Spatial balance measures (Voronoi and Modified Moran's I), Variance estimators (local mean, nearest neighbour and quasi bootstrap), DAS and the Spatial stratification algorithm.

This vignette is divided into the following sections:

## Simple Random Sampling (SRS)

### uc511::SRS()

This function invokes __base::sample()__ to draw a random sample using a user specified random seed. 

The following parameters are supported:

+ __seed__ - The random seed to be used to draw the current sample. The default is 42.
+ __total_rows__ - The total number of rows in our input dataset.
+ __sample_size__ - The number of rows wanted in our random sample.

All parameter values must be numeric and have values greater than zero, __sample_size__ must be less than __total_rows__.

#### uc511::SRS() code example

Function __uc511::SRS()__ returns a list of size __sample_size__.

```{r}
# Create a random sample with a seed of 99.
rand_samp <- uc511::SRS(seed = 99, total_rows = 100, sample_size = 20)
rand_samp
```

## Balanced acceptance sampling (BAS)

#### Load Shape File

##### Load Shape File
##### Select shape file from datafile

#### Specify Parameters

##### Sample Size
##### Random Seed Options
##### Use st_transform
##### Minimum Distance

#### Generate Sample

##### Generate BAS Sample
##### Download BAS Sample

#### Visualize Sample

##### Visualize BAS Sample on Map
##### Visualize BAS Sample Points

## Halton Iterative Partitioning (HIP)

Halton iterative partitioning (HIP) extends Basic acceptance sampling (BAS) to point resources. It partitions the resource into $B ≥ n$ boxes that have the same nested structure as in BAS, but different sizes. These boxes are then uniquely numbered using a random-start Halton sequence of length $B$. The HIP sample is obtained by randomly drawing one point from each of the boxes numbered $1, 2, . . . , n$.

This function invokes __uc511::cppRSHalton_br()__ to generate the Halton points.

### uc511::HIP()

Describe HIP() and its parameters.

The __uc511::HIP()__ function accepts the following parameters:

#### population

The __population__ parameter is required (there is no default), this is a list of points comprising the population from which the sample will be drawn.

#### n

The __n__ parameter is the number of points to draw from the population. The default value is __20__.

#### iterations

The __iterations__ parameter is the number of levels of partitioning required. The default value is __7__.

### uc511::HIP() output

The __uc511::HIP__ function returns the following output in the form of a list of variables. The variable names and what they contain are as follows:

+ __sampleI__ - indices of the selected sample points from the population.
+ __popIndex__ - __*what are these used for?*__
+ __Order__ - __*what are these used for?*__
+ __HaltonIndex__ - The Halton indices for all the population points.


### uc511::HIP() code example

The following is sample code to demonstrate the use of the __uc511::HIP()__ function. Here we are generating $20$ points from a population of $5,000$ (random) points with $7$ levels of partitioning ($4$ in the first dimension and $3$ in the second) to give $2^4 = 32$ by $3^3 = 27$ boxes.

```{r}
# set random seed
set.seed(1)

# define HIP parameters.
pop <- matrix(runif(5000*2), nrow = 5000, ncol = 2)
n <- 20
its <- 7

# generate HIP sample.
result <- uc511::HIP(population = pop, n = n, iterations = its)

# sampleI - indices of the selected samples from the population.
sampleI <- result$sampleI
sampleI

# popIndex
popIndex <- result$popIndex
head(popIndex)

# Order
Order <- result$Order
head(Order)

# HaltonIndex
HaltonIndex <- result$HaltonIndex
table(HaltonIndex)
```


## Halton Frames (HF)

A Halton frame is a useful grid structure, it is a collection of Halton boxes that intersect a resource.

### uc511::HaltonFrame()

This function invokes __uc511::cppRSHalton_br()__ to generate Halton points. 

The following parameters are supported:

+ __J__ - The ??????. The default is $c(3, 2)$.
+ __bases__ - The co-prime base for the Halton sequence. The default is $c(2, 3)$.
+ __n__ - The number of Halton points to generate. The default is ($bases[1]^{J[1]} * bases[2]^{J[2]}$).

All parameter values must be numeric and have values greater than zero. The __J__ and __bases__ parameters must each be a list of length two.

#### uc511::HaltonFrame() code example

Function __uc511::HaltonFrame()__ returns a list of variables, the variables have the following names:

+ __halton_frame__ the generated Halton frame.
+ __halton_sequence__ the generated Halton sequence.
+ __Z__ something.
+ __halton_seq_dev__ the $x$ and $y$ axis values that can be used when plotting to sub-divide the Halton frame.

```{r}
# Define parameters.
J <- c(3, 2)
bases <- c(2, 3)
n <- (bases[1]^J[1]) * (bases[2]^J[2])

# generate Halton frame.
hf_ <- uc511::HaltonFrame(n = n, J = J, bases = bases)

# retrieve the Halton sequence.
halton_seq <- hf_$halton_seq
head(halton_seq)

# retrieve the x-axis division values.
halton_seq_x_divisions <- hf_$halton_seq_div[1]
halton_seq_x_divisions

# retrieve the y-axis division values.
halton_seq_y_divisions <- hf_$halton_seq_div[2]
halton_seq_y_divisions

# retrieve the Z values.
z <- hf_$Z
head(z)

# retrieve the Halton frame.
halton_frame <- hf_$halton_frame
head(halton_frame)
```


# The uc511 Shiny App for Spatially balanced sampling.

Part of the uc511 package is the uc511 Shiny App, this provides a user interface to the uc511 package. Currently the following options are supported:

+ __SRS__ Simple random sampling using the __uc511::SRS()__ function.
+ __BAS__ Balanced acceptance sampling using the __uc511::BAS()__ function.
+ __HIP__ Balanced acceptance sampling using the __uc511::HIP()__ function.

The following sections detail how to use the supported features in the uc511 Shiny App.

## uc511::SRS() from the uc511 Shiny App

The __uc511__ __R__ package also includes a Shiny App, this has a menu option (SRS) to invoke Simple Random Sampling. The _SRS_ menu option supports the following sub-menu's, __Load Auxillary File__, __Specify Parameters__, __Generate SRS Sample__, __something__ and __something__. We will now look in turn at the options provided by each menu item. Each menu item will not enable until all relevant options on a previous menu item have been appropriately selected or provided.

### Load Auxillary File

Here the user selects the "Browse..." button to initiate the file selection dialog. The selection of a single CSV file is supported. The application assumes that the first row of the CSV contains column headings. Once a CSV file has been selected and loaded the __Specify Parameters__ menu item is enabled.

### Specify Parameters

With this menu item, the user is able to specify parameters that will be subsequently passed to the __uc511::SRS__ function. The __Specify Parameters__ menu item supports the following tab panels: __Sample source__, __Sample size__, __Random seed options__, __Select X and Y columns__, and __Projected coordinate system__. We will now look at each option in turn, to see what they do and how they are used.

#### Sample source

The __Sample source__ option controls what type of sample is generated from the auxillary file. At the time of writing only two options are supported:

* __Sample from Auxillary File__ This is the default. When this option is selected, the user wants to create a random sample from the previously loaded auxillary file.

* __Create Master Frame and take sample__ When this option is selected, a Master Frame (need to define this) is created from the auxillary file and then a random sample is then created from the Master Frame.

#### Sample size

The __Sample size__ option controls the size of the generated sample. At the time of writing only two options are supported:

* __Sample size as a percentage__ This is the default. If this option is selected, a slider is displayed. Use the slider to select the percentage of the auxillary file to be used as the sample. The corresponding sample size is displayed as an integer in a textbox to the right of the slider.

* __User specified sample size__ (Rename this option in the app and update here!). If this option is selected, the user is able to specify a specific sample size in a text inputbox. A sample size in the range one upto the number of rows in the auxillary file is supported.

The __Generate Sample__ menu item will only enable once the application user has specified a sample size.


#### Random seed options

The __Random seed options__ option controls how the random seed is set. At the time of writing only two options are supported:

* __Set seed randomly__ This is the default. If this option is selected, the shiny app. will generate a random seed in the range __1-10,000__ (verify this!).

* __User specified seed__ If this option is selected, a textinput box is displayed, which gives the user the opportunity to specify their own seed. Any valid integer is acceptable (check if I validate this!). A default of __42__ is provided.

#### Select X and Y columns

The __Select X and Y columns__ option gives the user the chance to select the columns to use for both the X and Y axes on the __Visualize Sample__ menu option.

* __Select column to use for X__ From the drop down box select a single column name to use as the X-axis.

* __Select column to use for Y__ From the drop down box select a single column name to use as the Y-axis.

Both drop down boxes are populated using the column names obtained from the auxillary file.

#### Projected coordinate system

This parameter is only used when the __Visualize Sample__ option is selected. It provides the EPSG projected coordinate system to be used to create the __proj4string__ (need to explain context).

* __Use CRS EPSG transform__ Check the check-box to have the option to specify a CRS EPSG code value. The default value is 28992 (where is this?).

If the checkbox is not checked, no coordinate system transform will be performed.

### Generate SRS Sample

With this menu item, the user is able to generate the random sample using the parameters previously specified. The __Generate SRS Sample__ menu item supports the following tab panels: __Generate SRS sample__, __Download SRS sample__, __Download Master Sample Indices__ and __Download Master Sample__. We will now look at each option in turn, to see what they do and how they are used.

#### Generate SRS sample

To generate the SRS sample, click the __Generate SRS Sample__ button. The random seed used to generate the sample is displayed in a textbox to the right of screen. If the user wants to generate a new sample with a new seed, then they will have to change the seed from the __Random seed options__ panel and then generate a new sample. The generated sample is displayed in a datatable for ease of perusal.

#### Download SRS sample

Use this option if you want to save the generated sample to a CSV file. Click the __Download__ button to initiate the file save process through a standard file save dialog. The default file name generated by the application has the form:

* __dataset-srs-sample-yyyy-mm-dd-seed.csv__

Where __yyyy__ is the year, __mm__ the month, __dd__ the day and __seed__ the seed used to generate the sample. The user is of course free to change the file name and the location of where the file will be saved.

#### Download Master Sample Indices

Use this option if you want to save the master sample indices to a CSV file. Click the __Download__ button to initiate the file save process through a standard file save dialog. The default file name generated by the application has the form:

* __dataset-srs-master-sample-indices-yyyy-mm-dd-seed.csv__ (need to add seed in code)

Where __yyyy__ is the year, __mm__ the month, __dd__ the day and __seed__ the seed used to generate the sample. The user is of course free to change the file name and the location of where the file will be saved.

#### Download Master Sample

Use this option if you want to save the master sample to a CSV file. Click the __Download__ button to initiate the file save process through a standard file save dialog. The default file name generated by the application has the form:

* __dataset-srs-master-frame-yyyy-mm-dd-seed.csv__ (need to add seed in code)

Where __yyyy__ is the year, __mm__ the month, __dd__ the day and __seed__ the seed used to generate the sample. The user is of course free to change the file name and the location of where the file will be saved.

### Visualize Sample

With this menu item, the user is able to generate a Leaflet map, the map is generated automatically when the __Visualize Sample__ menu option is selected. 

For the map to be generated successfully, the user would have had to use the __Specify Parameters__ menu option to specify the column names to use as the __X__ and __Y__ axes, using the __Select X and Y columns__ option and have specified a CRS EPSG transform code using the __Projected coordinate system__ option.

By default, both the original data points (black) and the sample points (red) are displayed on the map. The leaflet map control [---] provides the user the ability to select which set of points are displayed.

The user can click on any point and a popup message box will display the latitude and logitude of that point.

#### Save Image

Clicking on the __Save Image__ button will initiate process through a standard file save dialog. The default image file name generated by the application has the form:

* __map-srs-sample-yyyy-mm-dd-seed.jpeg__ (need to add seed in code, check if default is jpeg)

What is being displayed on the map at the time the __Save Image__ button is clicked is what is saved in the image file. 


## uc511::HaltonFrame() from the uc511 Shiny App

The __uc511__ __R__ package also includes a Shiny App, this has a menu option (Halton Frame) used to create a sample of Halton Points. The _Halton Frame_ menu option supports the following sub-menu's, __Load Shape File__, __Specify Parameters__, __Generate Sample__, __Visualize Frame__ and __Visualize Sample???__. We will now look in turn at the options provided by each menu item. Each menu item will not enable until all relevant options on a previous menu item have been appropriately selected or provided.

### Load Shape File

### Specify Parameters

With this menu item, the user is able to specify parameters that will be subsequently passed to the __uc511::HaltonFrame__ function. The __Specify Parameters__ menu item supports the following tab panels: __Sample size__, __J__, __Bases__, __Random seed options__ and __Use st_transform__. We will now look at each option in turn, to see what they do and how they are used.

#### Sample size

The __Sample size__ option controls the size of the generated sample. At the time of writing only two options are supported:

* __Sample size as a percentage__ This is the default. If this option is selected, a slider is displayed. Use the slider to select the percentage of the auxillary file to be used as the sample. The corresponding sample size is displayed as an integer in a textbox to the right of the slider.

* __User specified sample size__ (Rename this option in the app and update here!). If this option is selected, the user is able to specify a specific sample size in a text inputbox. A sample size in the range one upto the number of rows in the auxillary file is supported.

The __Generate Sample__ menu item will only enable once the application user has specified a sample size.

#### J

The __J__ option controls the... The default value is __(3, 2)__. The __J__ parameter is a list of two values.

#### Bases

The __Bases__ option is used to specify the values for the co-prime base for the Halton sequence. The default value is __(2, 3)__. The __Bases__ parameter is a list of two values.

#### Random seed options

The __Random seed options__ option controls how the random seed is set. At the time of writing only two options are supported:

* __Set seed randomly__ This is the default. If this option is selected, the shiny app. will generate a random seed in the range __1-10,000__ (verify this!).

* __User specified seed__ If this option is selected, a textinput box is displayed, which gives the user the opportunity to specify their own seed. Any valid integer is acceptable (check if I validate this!). A default of __42__ is provided.

#### Use st_transform


### Generate Sample

With this menu item, the user is able to generate the random sample using the parameters previously specified. The __Generate Sample__ menu item supports the following tab panels: __Generate Halton Frame__ __Download Halton Frame Sample__. We will now look at each option in turn, to see what they do and how they are used.

#### Generate Halton Frame

To generate the SRS sample, click the __Generate SRS Sample__ button. The random seed used to generate the sample is displayed in a textbox to the right of screen. If the user wants to generate a new sample with a new seed, then they will have to change the seed from the __Random seed options__ panel and then generate a new sample. The generated sample is displayed in a datatable for ease of perusal.

#### Download Halton Frame sample

Use this option if you want to save the generated sample to a CSV file. Click the __Download__ button to initiate the file save process through a standard file save dialog. The default file name generated by the application has the form:

* __dataset-srs-sample-yyyy-mm-dd-seed.csv__

Where __yyyy__ is the year, __mm__ the month, __dd__ the day and __seed__ the seed used to generate the sample. The user is of course free to change the file name and the location of where the file will be saved.


### Visualize Sample

With this menu item, the user is able to generate a Leaflet map, the map is generated automatically when the __Visualize Sample__ menu option is selected. 

For the map to be generated successfully, the user would have had to use the __Specify Parameters__ menu option to specify the column names to use as the __X__ and __Y__ axes, using the __Select X and Y columns__ option and have specified a CRS EPSG transform code using the __Projected coordinate system__ option.

By default, both the original data points (black) and the sample points (red) are displayed on the map. The leaflet map control [---] provides the user the ability to select which set of points are displayed.

The user can click on any point and a popup message box will display the latitude and longitude of that point.

#### Save Image

Clicking on the __Save Image__ button will initiate process through a standard file save dialog. The default image file name generated by the application has the form:

* __map-srs-sample-yyyy-mm-dd-seed.jpeg__ (need to add seed in code, check if default is jpeg)

What is being displayed on the map at the time the __Save Image__ button is clicked is what is saved in the image file.



